%!TEX root = ../template.tex
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% abstract-en.tex
%% NOVA thesis document file
%%
%% Abstract in English([^%]*)
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\typeout{NT FILE abstract-en.tex}%

%Nowadays, it has become increasingly necessary to process and analyze large volumes of data. The use of tools such as dashboards for analysis and decision support in the context of interactive data visualization has become essential. Dashboards oriented towards interactive data visualization aim primarily to address the various analytical questions defined by their users. These questions are answered through the use of visualization techniques and interaction taxonomies to represent information clearly and objectively. While visualization techniques allow data to be represented visually, interaction taxonomies help users explore the presented information from different perspectives. However, the process of creating dashboards often requires an iterative approach, demanding constant communication between designers and stakeholders. This communication process becomes quite complex and is sometimes overlooked due to the lack of support tools at this prototyping stage. As a response to this gap, \gls{IVML} emerged as a proposal for documenting and structuring the different components that make up a dashboard. The goal of this dissertation is to develop a tool based on the \gls{IVML} language, capturing all stages of dashboard creation—from the definition of the different analytical scopes to the implementation and final documentation of the visualizations and interactions in the finished dashboard. This tool will facilitate communication among the stakeholders in the design process, streamlining the entire creation and prototyping workflow. The tool will be implemented through the development of different prototypes and meta-models, which will be tested and validated to ensure their effectiveness and utility. With this approach, the entire dashboard creation process is expected to be optimized, enabling better adaptation to user needs and greater efficiency in the implementation of the proposed solutions.

The growing need to process and analyze large volumes of data has driven the emergence of analytical dashboards and tools that support their creation and use. These dashboards, oriented towards interactive data visualization, are designed to address the various analytical questions defined by their users. Such questions are answered through the use of visualization and interaction techniques that present information in a clear and objective way. While visualization techniques allow data to be represented visually, interaction helps users explore the information from different perspectives. However, the process of creating dashboards requires an iterative approach that demands constant communication among all participants involved. This communication process often becomes complex and is sometimes neglected due to the lack of supporting tools during the prototyping stage. As a response to this issue, IVML emerged as an initial proposal for documenting and structuring the different components that compose a dashboard. The objective of this dissertation is the development of Dashboard Designer, a web-based tool that uses the IVML language as its foundation, capturing all stages of dashboard creation, from defining the different analytical components that determine their objectives and purposes to the implementation and final documentation of the visualizations and interactions included in the finished dashboard. This tool will facilitate communication among all participants in the design process, making the creation and prototyping stage easier. The tool will be implemented through the development of different prototypes and models, which will be tested and validated to ensure their effectiveness and usefulness. With this approach, it is expected that the entire dashboard creation process will be optimized, enabling better adaptation to user needs and greater efficiency in implementing the proposed solutions.

% Palavras-chave do resumo em Inglês
% \begin{keywords}
% Keyword 1, Keyword 2, Keyword 3, Keyword 4, Keyword 5, Keyword 6, Keyword 7, Keyword 8, Keyword 9
% \end{keywords}
\keywords{
  Interactive Data Visualization \and
  Design Process \and
  Analytical Dashboards \and
  Dashboard Designer \and
  Web-based Tool \and 
  IVML
}
